---
title: "PA1_template"
author: "Muhammad Arshad khan"
date: "December 11, 2016"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1-Loading Data

```{r myfile, echo=TRUE}
myfile<-read.csv("activity.csv",header=TRUE, sep=",")
head(myfile)
```

## 2-What is mean total number of steps taken per day?
```{r , echo=TRUE}
vtot_Steps <-aggregate(steps~date,data=myfile,FUN=sum)
head(vtot_Steps)
```

## 3-Make a histogram of the total number of steps taken each day

```{r , echo=TRUE,error=TRUE,fig.width=7, fig.height=4}
library(ggplot2)
g <-ggplot(vtot_Steps,aes(x=vtot_Steps$date,y=vtot_Steps$steps))+geom_histogram(stat="identity")
g
```

## 4-Calculate and report the mean and median of the total number of steps taken per day

```{r , echo=TRUE,error=TRUE,fig.width=7, fig.height=4}
myFun <- function(x) {
  c(mean = mean(x), median = median(x)
  )
}
mymean<- aggregate(steps ~date , data=myfile[!is.null(myfile$steps),],FUN=myFun)
head(mymean)
```

## 5-Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r , echo=TRUE,error=TRUE,fig.width=7, fig.height=4}
mymean<- aggregate(steps ~ interval , data=myfile[!is.null(myfile$steps),],FUN=mean)
plot(steps ~ interval,data=mymean,type="l")
```

## 6-Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r , echo=TRUE,error=TRUE,fig.width=7, fig.height=4}
maxinterval <-subset(mymean,steps ==max(mymean$steps))
print (maxinterval[,-2])
```

## 7-Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r , echo=TRUE,error=TRUE,fig.width=7, fig.height=4}
total_null<-sum(is.na(myfile))
print (total_null)
```
## 8-Devise a strategy for filling in all of the missing values in the dataset.
```{r , echo=TRUE,error=TRUE,fig.width=7, fig.height=4}
myfile[is.na(myfile)] <- 0
print (which (is.na(myfile)))
```
## 9-Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r , echo=TRUE,error=TRUE,fig.width=7, fig.height=4}
myfile <-read.csv("activity.csv",header = TRUE)
mynewfile <-myfile
mynewfile[is.na(mynewfile)] <- 0
```
## 10- Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r , echo=TRUE,error=TRUE,fig.width=7, fig.height=4}
vtot_Steps <-aggregate(steps~date,data=mynewfile,FUN=sum)
library(ggplot2)
g <-ggplot(vtot_Steps,aes(x=vtot_Steps$date,y=vtot_Steps$steps)) +geom_histogram(stat="identity")
g   
```
## 11- Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r , echo=TRUE,error=TRUE,fig.width=7, fig.height=4}
myweekend <- function(x) {
  if( x=="Friday" | x=="Saturday"){
    print("Weekend")} 
  else {
    print("Weekday")
  }
}

mynewfile[,"weekday"]<-sapply(weekdays(as.Date(mynewfile$date,'%Y-%m-%d')),FUN=myweekend)
head(mynewfile)

```
## 12- Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r , echo=TRUE,error=TRUE,fig.width=25, fig.height=25}

mymeanweekdays<- aggregate(steps ~ interval , data=subset(mynewfile,mynewfile$weekday =="Weekday"),FUN=mean)
mymeanweekend<- aggregate(steps ~ interval , data=subset(mynewfile,mynewfile$weekday =="Weekend"),FUN=mean)
par(mfrow=c(2,1))
plot(steps ~ interval,data=mymeanweekdays,type="l",main="Weekdays")
plot(steps ~ interval,data=mymeanweekend,type="l",main="Weekend")

```














